---
title: "homework 5"
author: "Bogdan Abaev"
date: "2/26/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
```
Option 2

Question 1:
regular expression "\^[Ss](.*)(t+)(.+)(t+)" will match a string that starts with lower-case or upper-case "s", followed by any (including 0) occurence of any character, followed by at least one t, followed by at least one occurence of any character, followed by at least one occurence of t. “scuttlebutt”, “Stetson”, and “Scattter” fit this pattern, since each starts with an "s", and then each has at least two "t"'s with some characters in between them (this requirement comes from "(.+)" part of the expression). “Scatter" does have 2 occurences of "t", however, there is no character between them. Thus, the second "t" in "Scatter" is matched with "(.+)" part of the expression, and there is no "t" in the string to match with the second "(t+)" in the regular expression. This is why "Scatter" is not matched with "\^[Ss](.*)(t+)(.+)(t+)".
```{r,include=TRUE}
# question 2
emails = c(
"simple@example.com",
"johnsmith@email.gov",
"marie.curie@college.edu",
"very_common@example.com",
"a.little.lengthy.but.ok@dept.example.com",
"bad.email.because+symbol@example.com",
"not_good@email.address",
"this.email.is.fake@gmail.xcom")

# list of valid domains. One can add more domains to the list if needed
#(there are more than 1500 valid domains, so including them all wouldn't make
#sense for the assignment)
doms = c("com", "uk", "gov", "edu", "fr")

#find and return the domain of the email
# by separating string by . and taking the last substring
domain = function(x){
  x = unlist(strsplit(x,split='.', fixed = TRUE))
  return(x[length(x)])
}

# check if the domain of the email is in the list of
# valid domains
has_good_domain = function(x){
  domain = domain(x)
  return (domain %in% doms) 
}

# check if the email has only letters and numbers
letters_nums_only <- function(x) !grepl("[^[:alnum:] ]", x)


# drop the non-alphanumeric characters that are allowed
# in a valid email from the string
drop_fine_chars = function(x) {
  a = gsub("_", "", x, fixed=TRUE)
  a = gsub(".","",a, fixed=TRUE)
  a = gsub("-","",a, fixed=TRUE)
  a = gsub("@","",a, fixed=TRUE)
  return(a)
}
#check if the email is valid by checking if
#the domain is valid, and then dropping all valid
# non-alphanumeric characters and seeing if there are
# any other non-alphanumeric characters (that are not allowed)
valid_email = function(x){
  good_domain = has_good_domain(x)
  y = drop_fine_chars(x)
  return (letters_nums_only(y)  && good_domain)
}

print(mapply(valid_email, emails))
```


