---
title: "Homework 1"
author: "Bogdan Abaev"
date: "1/23/2019"
output:
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```
<center>
Option 2 for homework
</center>
```{r}
#Question 1
Dat = read.csv("https://remiller1450.github.io/data/CollegeData.csv")
#Question 4
table(Dat$REGION,Dat$LOCALE)
```

The table above shows the two way relationship between the locale and the region variables. The top row represents the locales, and the left most column represents regions.
Thus, every column Y from the table presents data about the number of schools from regions 1 ... 8 in Y locale, and every row X shows how many schools from X region fall into each of the locales.
We can observe that most colleges are located in the 5th region all over different locales, with most colleges being present in locale 32.

```{r, include=TRUE}
#Question 7
# vector - a valid non-empty vector of numeric values
# percent - a decimal between 0 and 1
trimmedmean = function(vector, percent){
  #sort the vector
  sorted = sort(vector)
  # find the bottom and top cut-offs
  leftPerc = quantile(sorted, percent)
  rightPerc = quantile(sorted, 1-percent)
  #trimm
  trimmed = vector[vector < rightPerc & vector > leftPerc]
  return (mean(trimmed))
}

trimmedmean(Dat$AVGFACSAL, 0.05)
```




