---
title: "HW_4_Bogdan_Abaev"
author: "Bogdan Abaev"
date: "2/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("Lahman")
library(Lahman)
library(dplyr)
```

Option 2
```{r, include = TRUE}
# A
player_df = select(Master, playerID, nameFirst,nameLast)

# B
no_salary = anti_join(x=player_df,y=Salaries,by="playerID")
head(no_salary)
tail(no_salary)
dim(no_salary)
```


```{r, include=TRUE}
#c
dont_appear = anti_join(no_salary, Appearances,by="playerID")
dim(dont_appear)
```
We can see that there are 193 players who do not have a salary that are not present in Appearances table. Thus, 
almost all players (>13500) from no_salary table did participate in games of their teams.

```{r}
#d
batting_df = select(Batting, playerID, AB)
no_salary_batting = inner_join(no_salary, batting_df, "playerID")
group = group_by(no_salary_batting,playerID)
non_salary_atbats=summarise(group, ab = sum(AB, na.rm = TRUE))
head(non_salary_atbats)
tail(non_salary_atbats)
dim(non_salary_atbats)
dim(filter(non_salary_atbats, ab < 100))

```

Here we can see the table with players who do not hava a salary next to their total batting. Almost everyone from the 
non-salary table is present in this table as well, therefore, in general, players did play in games they appeared in.
However, one can note that more than a half of players from the table have a fairly low total batting, so this can be one of possible explanations to why there's no data on their salaries.







